---
title: "Arnab's Flexdashboard"
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: fill
    orientation: rows
    theme: united
    social: ["twitter", "facebook", "menu"]
---

```{r setup, include=FALSE}
library(flexdashboard)
library(dplyr)
library(ggplot2)
library(plotly)
library(tidyr)
library(forcats)
library(highcharter)
library(knitr)
library(DT)
library(openintro)
library(ggvis)
library(data.table)
library(viridis)
library(crosstalk)
```

```{r}
data <- read.csv("C:/Users/ararn/Documents/insurance.csv")
```

```{r}
mycolors <- c("blue", "#FFC125", "darkgreen", "darkorange", "pink", "tomato")
```

```{r include=FALSE}
# Create a custom theme for the plots. 
custom_theme <- hc_theme(
  colors = c('#5CACEE', 'green', 'red'),
  chart = list(
         backgroundColor = '#FAFAFA', 
         plotBorderColor = "black"),
  xAxis = list(
         gridLineColor = "E5E5E5", 
         labels = list(style = list(color = "#333333")), 
         lineColor = "#E5E5E5", 
         minorGridLineColor = "#E5E5E5", 
         tickColor = "#E5E5E5", 
         title = list(style = list(color = "#333333"))), 
  yAxis = list(
         gridLineColor = "#E5E5E5", 
         labels = list(style = list(color = "#333333")), 
         lineColor = "#E5E5E5", 
         minorGridLineColor = "#E5E5E5", 
         tickColor = "#E5E5E5", 
         tickWidth = 1, 
         title = list(style = list(color = "#333333"))),   
  title = list(style = list(color = '#333333', fontFamily = "Lato")),
  subtitle = list(style = list(color = '#666666', fontFamily = "Lato")),
  legend = list(
         itemStyle = list(color = "#333333"), 
         itemHoverStyle = list(color = "#FFF"), 
         itemHiddenStyle = list(color = "#606063")), 
  credits = list(style = list(color = "#666")),
  itemHoverStyle = list(color = 'gray'))
```

Interactive Data Visualization
=====================================

Row
-----------------------------------------------------------------------

### Medical Analysis of US

```{r}
valueBox(paste("Medical Analysis"),
         color = "warning")
```

### Medical Problems in US

```{r}
valueBox(length(data$State),
         icon = "fa-user")
```

### **BMI**

```{r}
gauge(round(mean(data$bmi),
            digits = 2),
            min = 0,
            max = 55,
            gaugeSectors(warning = c(0, 18),
                         success = c(18, 26),
                         danger = c(26, 55),
                         colors = c("yellow", "green", "red")))
```

### Texas

```{r}
valueBox(sum(data$State == "TX"),
         icon = 'fa-building')
```

### California

```{r}
valueBox(sum(data$State == "CA"),
         icon = 'fa-building')
```

### Florida

```{r}
valueBox(sum(data$State == "FL"),
         icon = 'fa-building')
```

### Arizona

```{r}
valueBox(sum(data$State == "AZ"),
         icon = 'fa-building')
```

Row
-------------------------------

### Meical Issues By State

```{r}
p1 <- data %>%
         group_by(State) %>%
         summarise(count = n()) %>%
         plot_ly(x = ~State,
                 y = ~count,
                 color = "blue",
                 type = 'bar') %>%
layout(xaxis = list(title = "Medical Issues By State"),
yaxis = list(title = 'Count'))
p1
```

### Pie Chart of Top States

```{r}
p2 <- data %>%
         group_by(State) %>%
         summarise(count = n()) %>%
         filter(count>50) %>%
         plot_ly(labels = ~State,
                 values = ~count,
                 marker = list(colors = mycolors)) %>%
         add_pie(hole = 0.2) %>%
         layout(xaxis = list(zeroline = F,
                             showline = F,
                             showticklabels = F,
                             showgrid = F),
                yaxis = list(zeroline = F,
                             showline = F,
                             showticklabels=F,
                             showgrid=F))
p2
```

### Age Vs BMI

```{r}
p3 <- plot_ly(data,
              x = ~age,
              y = ~bmi,
              text = paste("Age:", data$age,
                           "BMI:",
                           data$bmi),
              type = "bar") %>%
         layout(xaxis = list(title="Age"),
                yaxis = list(title = "BMI"))
p3
```

Row
------------------------------------
### Scatter Plot of Month Vs Charges

```{r}
p4 <- plot_ly(data, x=~age) %>%
         add_markers(y = ~charges,
                     text = ~paste("Cost: ", charges),
                     showlegend = F) %>%
         add_lines(y = ~fitted(loess(charges ~ age)),
                   name = "Average spending",
                   color = I("#FFC125"),
                   showlegend = T,
                   line = list(width=5)) %>%
         layout(xaxis = list(title = "Age"),
                yaxis = list(title = "Charges"))
p4
```

### Box Plot of Regions

```{r}
p5 <- data %>%
  ggplot( aes(region, charges, size = charges, color=region)) +
  geom_point() +
  theme_bw()
p5
```

### Box Plot of Medical Charges by Smoking Status

```{r}
ggplot(data = data,aes(smoker,charges)) + geom_boxplot(fill = c(2:3)) +
  theme_classic()
```

Map
========================================

### Map

```{r}
insurance <- data %>%
         group_by(State) %>%
         summarize(total = n())

insurance$State <- abbr2state(insurance$State)

highchart() %>%
         hc_title(text = "Total insurances in US") %>%
         hc_subtitle(text = "Source: insurance.csv") %>%
         hc_add_series_map(usgeojson, insurance,
                           name = "State",
                           value = "total",
                           joinBy = c("woename", "State")) %>%
         hc_mapNavigation(enabled = T)
```


Data Table
========================================

```{r}
datatable(data,
          caption = "Medical Cost Data",
          rownames = T,
          filter = "top",
          options = list(pageLength = 30))
```

Total Charges
========================================================================

Column {.tabset .tabset-fade data-width=400 .colored }
-----------------------------------------------------------------------

### By state {.no-padding}
```{r fig.height=4.55}
data <- as.data.table(data)
data[,count := .N,by="State"]

# Create tibble for overall charges by state
State_bar <- data %>%
  select(State, charges, count) %>%
  group_by(State) %>%
  summarise(charges_capita = round((sum(charges)/sum(count)), 2)) %>%
  arrange(desc(charges_capita))

# Create interactive bar plot
highchart() %>%
    hc_add_series(State_bar, hcaes(x = State, y = charges_capita, color = charges_capita), type = "bar")  %>% 
    hc_tooltip(borderWidth = 1.5, 
               pointFormat = paste("Charges: <b>{point.y}</b>")) %>%
    hc_legend(enabled = FALSE) %>%
    hc_title(text = "Charges by State") %>% 
    hc_xAxis(categories = State_bar$State, 
             labels = list(step = 1),
             min = 0, max = 25,
             scrollbar = list(enabled = TRUE)) %>%
    hc_yAxis(title = list(text = "Total Charges")) %>%
    hc_plotOptions(bar = list(stacking = "normal", 
                              pointPadding = 0, groupPadding = 0, borderWidth = 0.5)) %>%
  hc_add_theme(custom_theme)
```

### By Gender and Age {.no-padding}
```{r}
# Create tibble for charges by age and sex. 
sex_color <- c("pink", "tomato")
age_bar_sex <- data %>%
  select(age, sex, count, charges) %>%
  group_by(age, sex) %>%
  summarise(charges_capita = round((sum(charges)/sum(count)), 2))

# Create bar chart of charges by sex.
highchart() %>%
    hc_add_series(age_bar_sex, hcaes(x = age, y = charges_capita, group = sex), type = "bar", color = sex_color)  %>% 
    hc_tooltip(borderWidth = 1.5, pointFormat = paste("Gender: <b>{point.sex} ({point.percentage:.1f}%)</b> <br> Round up charges: <b>{point.y}</b>")) %>%
    hc_legend(enabled = TRUE, colorByPoint = TRUE) %>%
    hc_title(text = "Charges by age and gender") %>% 
    hc_plotOptions(bar = list(stacking = "percent", 
                              pointPadding = 0, groupPadding = 0, borderWidth = 0.4)) %>% 
    hc_add_theme(custom_theme)
```

Dynamic Search {data-orientation=columns}
=======================================================================

Column {data-width=275}
-----------------------------------------------------------------------

### Filters {.no-title .colored }

**Pick/Select filters here:**

```{r}
# Create tibble for our line plot. 
State_bmi_tibble <- data %>%
  select(State, bmi, charges, count, smoker) %>%
  group_by(State, bmi, smoker) %>%
  summarise(charges_capita = round((sum(charges)), 2))

# Create shared data that will be used to link filters, data table, and line plot. 
shared_data <- SharedData$new(State_bmi_tibble, group = 'hello')

# Create filter for year and country. These filters will adjust the DT datatable and PLOTLY plot. 
filter_slider("Bmi", "Bmi", shared_data, ~bmi, step = 1)
filter_select("State", "State", shared_data, ~State, allLevels = TRUE, multiple = TRUE)
filter_select("Smoker", "Smoker", shared_data, ~smoker, allLevels = TRUE, multiple = TRUE)
```

***

```{r}
# Create datatable. 
datatable(shared_data,
          rownames = FALSE,
          colnames = c('State', 'Bmi', 'Smoker', 'Charges'),
          class = 'cell-border stripe',
          width = '100%',
          extensions = "Scroller",
          options=list(deferRender = FALSE, 
                       scrollY = 280, 
                       scrollCollapse = TRUE,
                       scroller = TRUE,
                       dom = 't'))
```

Column {data-width=725}
-----------------------------------------------------------------------

### Charges by state {.no-title .colored }
```{r}
# Set a random seed. We will do this so that we can reproduce the random sample of colors we will use for our next graph. 
set.seed(80085)

# Create line graph. 
plot_ly(shared_data, x = ~bmi, y = ~charges_capita, 
       color = ~State, colors = sample(colours(), 120),
        mode = 'lines',
       hoverinfo = 'text', text = ~paste("State: ", State, '<br>Bmi: ', bmi, "<br>Charges: ", charges_capita, "<br>Smoker: ", smoker)) %>%
       layout(showlegend = TRUE,
              title = "Charges by State on different BMIs",
              xaxis = list(title = "Bmi"),
              yaxis = list(title = "Charges")) %>%
       layout(plot_bgcolor = 'transparent') %>% 
       layout(paper_bgcolor = 'transparent') %>% 
       add_markers() %>% 
       highlight("plotly_click")
```

Summary 
===========================================

---------------------------

Report

* The dataset is regarding Medical Cost of different individuals of United States with 1338 rows.

* The dataset is downloaded from https://www.kaggle.com/mirichoi0218/insurance .

* The dataset has different features like Charges(total spending), Age, Sex, Smoker or Non Smoker, BMI(body mass index), Childrens.

* The flexdashboard contains various interactive visualizations. 

The main purpose behind the visualizations was to learn on what factors does medical charges increases?

The medical charges increases with ->

1. Increase in age.
2. Increase in BMI.
3. If he/she is a smoker.

The above answers are depicted in different visualization plots shown in the previous tabs.

Row 
-----------------------------------

### Maximum Spent

```{r}
valueBox(max(data$charges),
         icon = "fa-user" )
```

### Average Spent
```{r}
valueBox(round(mean(data$charges),
               digits = 2),
         icon = "fa-area-chart")
```

### Average BMI

```{r}
valueBox(round(mean(data$bmi), digits = 2),
         icon = "fa-area-chart")
```

Row 
-----------------------------------

THANK YOU

About Report
========================================

STATISTICS FOR DATA SCIENCE

The Interactive Visualisation report:

Created by: Arnab Roy

Date: 29th April 2020

Course: Masters in Data Science

C00246500

Institute of Carlow, Carlow